% This is a template for Ph.D. dissertations in the UCI format.

% All fonts, including those for sub- and superscripts, must be 10 points or larger.
% Recommended sizes are 14-point for chapter headings, 12-point for the main body of text
% and figure/table titles, and 10-point for footnotes, sub- and super-scripts, and text in
% figures and tables.
\documentclass[12pt,fleqn]{ucithesis}

\usepackage{amsmath}
\usepackage{array}
\usepackage{bm}
\usepackage{boxedminipage}
\usepackage{graphicx}
%\usepackage{natbib}
%http://merkel.zoneo.net/Latex/natbib.php
\usepackage[numbers]{natbib}
\usepackage{path}
\usepackage{psfrag}
\usepackage{relsize}
%\usepackage{subfigure}
%\usepackage{subfig}
\usepackage{todonotes}
\usepackage{bytefield}
\usepackage{url}
\usepackage{verbatim}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}

% plainpages=false fixes the "duplicate ignored" error with page counters
% Set pdfborder to 0 0 0 to disable colored borders around PDF hyperlinks
\usepackage[plainpages=false,pdfborder={0 0 0}]{hyperref}



\graphicspath{{figures/}}

% Uncomment the following line to enable Unicode support. This will allow you
% to enter non-ASCII characters (such as accented characters) directly without
% having to use LaTeX's awkward escape syntax (e.g., \'{e})
% NOTE: You may have to install the ucs.sty package for this to work. See:
% http://www.unruh.de/DniQ/latex/unicode/
% \usepackage[utf8x]{inputenc}

\begin{document}


\thesistitle
{
  Improving the Architecture of Calico
}

\degreename{Master of Science}

% Use the wording given in the official list of degrees awarded by UCI:
% http://www.rgs.uci.edu/grad/academic/degrees_offered.htm
\degreefield{Information and Computer Science}

% Your name as it appears on official UCI records.
\authorname{Mitchell Ryan Dempsey}

% Use the full name of each committee member.
\committeechair{Professor Andr\'{e} van der Hoek}
\othercommitteemembers
{
  Professor James A. Jones\\
  Professor Richard N. Taylor
}

\degreeyear{2012}

\copyrightdeclaration
{
  {\copyright} {\Degreeyear} \Authorname
}

% If you have previously published parts of your manuscript, you must list the
% copyright holders; see Section 3.2 of the UCI Thesis and Dissertation Manual.
% Otherwise, this section may be omitted.
% \prepublishedcopyrightdeclaration
% {
%   Chapter 4 {\copyright} 2003 Springer-Verlag \\
%   Portion of Chapter 5 {\copyright} 1999 John Wiley \& Sons, Inc. \\
%   All other materials {\copyright} {\Degreeyear} \Authorname
% }

% The dedication page is optional.
% \dedications
% {
%   To my parents...
% }

\acknowledgments
{
  I would like to thank my advisor, Professor Andr\'{e} van der Hoek. Without his help and guidance my thesis would not be possible.
  I want to thank the members of the Software Design and Collaboration Lab: Nick Mangano, Nicolas Lopez, Gerald Bortis, Alex Baker, Tiago Proenca, and Nilmax Moura. I truly enjoyed my time spent with them, and greatly appreciate the help they have provided over the years.
}

% max 250 words
\thesisabstract
{
  Calico is a software design sketching tool that aims to support software designers by helping them work through a design problem by sketching potential solutions. 
  Calico was designed to be used with an interactive whiteboard so that it can replicate the feel of a traditional whiteboard while at the same time enabling advanced functionality. 

  There were three main problems with the original version of Calico: 
  (1) distributed designers were not well supported, 
  (2) collaboration was not natively supported,
  and (3) it was very difficult to add new functionality.
  This thesis examines these problems and presents various solutions:
  (1) creating a performance-based network architecture,
  (2) centering Calico around collaboration,
  and (3) creating a plugin framework to easily add new functionality.

  Prior to this work, Calico was unable to meet the demands of designers in the field. By resolving these problems, Calico was able to be deployed both in-house in our design lab, as well as in Informatics 122 (a software design course).
}

\preliminarypages

\input{./chapters/introduction.tex}

\input{./chapters/background.tex}

\input{./chapters/calico.tex}

\input{./chapters/objectives.tex}

\input{./chapters/high_level_architecture.tex}

% Implementation (broken into sections)
\input{./chapters/implementation.tex}
\input{./chapters/implementation/networking.tex}
\input{./chapters/implementation/object_storage.tex}
\input{./chapters/implementation/object_controllers.tex}
\input{./chapters/implementation/plugin_framework.tex}
\input{./chapters/implementation/input_handling.tex}
\input{./chapters/implementation/admin_ui.tex}

\input{./chapters/conclusion.tex}

% These commands fix an odd problem in which the bibliography line
% of the Table of Contents shows the wrong page number.
\clearpage
\phantomsection

% plain
%\bibliographystyle{abbrv} % DEFAULT
\bibliographystyle{acm}
%\bibliographystyle{abbrvnat}
\bibliography{thesis}

\appendix
\input{./appendices/calico_commands.tex}

\end{document}
